Transmembrane helices are also an overlooked source of major histocompatibility complex class II epitopes and evolutionary more conserved than expected by chance

### 2021-04-22 bbbq_article.tex
Changes Geert and Maxim

Although our data show that 
TMH-derived epitopes are presented in MHC-I and MHC-II, 
the molecular mechanisms of how integral membrane proteins are processed 
for MHC presentation are largely unknown (\cite{bianchi2017}). 
Most prominently, the fundamental principles of 
how TMHs are extracted from the hydrophobic lipid environment 
into the aqueous vacuolar lumen, 
and their prior or subsequent proteolytic processing are unresolved, as we discussed previously for MHC-I presentation \cite{bianchi2017}. 
\geert{
  Hier toevoegen wat probleem is MHC-II. Maxim had goed idee voor intraluminal vesicles in MHC-II loading.
}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Experimental evidence for MHC presentation of TMH-derived epitopes}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\geert{Deze gehele subsection kan wat mij betreft wel weg (of naar de 'minor discussion' in de supplement). Het is deels een herhaling van resultaten (1e en 3e paragraaf). DRiPs in paragraaf 2 kan echt niet alles verklaren, want MHC-II is opname van heel antigen. In parafraag 4, zal de methodological bias zal alleen maar resulteren in het onderschatten van presentatie van TMHs, en is dus ook niet echt relevant hier.}
% \paragraph{Over-presentation in vivo in MHC-I and MHC-II is robust
From reanalysis of peptide elution studies, we show that 1.4\% of epitopes that are naturally presented on MHC-I 
can be mapped to stem from a TMH. 
We also find that MHC-II presents epitopes derived from TMHs, 
as we find that 3.9-4.0\% of sequenced epitopes
can be uniquely traced onto a human reference proteome (also shown in table \ref{tab:elution}). \geert{zie eerdere vraag. Zijn deze percentages van totale peptides, of van humane peptides?}

>>>>

Although our data show that 
TMH-derived epitopes are presented in MHC-I and MHC-II, 
the molecular mechanisms of how integral membrane proteins are processed 
for MHC presentation are largely unknown (\cite{bianchi2017}). 
Most prominently, the fundamental principles of 
how TMHs are extracted from their hydrophobic lipid environments 
into the aqueous vacuolar lumen, 
and their prior or subsequent proteolytic processing are unresolved.

A first possibility is that the extraction of TMPs from the membrane is mediated by the the ER-associated degradation (ERAD) machinery. For MHC class I (MHC-I) antigen presentation of soluble proteins, the loading of the epitope primarily occurs at the endoplasmic reticulum (ER). The chaperones tapasin (TAPBP), ERp57 (PDIA3), and calreticulin (CALR) (Rock et al., 2016) first assemble and stabilize the heavy and light chains of MHC-I. Later, this complex binds to the transporter associated with antigen processing (TAP) leading to the formation of the so-called peptide-loading complex (PLC). The PLC drives import of peptides into the ER and mediates their subsequent loading into the peptide-binding groove of MHC-I (Blees et al., 2017). Membrane proteins first will have to be extracted from the membrane before they become amenable to this MHC-I loading by the PLC. In the ER, this process can be orchestrated by the ERAD machinery, consisting of several chaperones that recognize TMPs, ubiquitinate them, and extract them from the ER membrane into the cytosol (retrotranslocation) for proteasomal degradation (Preston GM, 2017, Meusser B 2005). Similar to the peptides generated from soluble proteins, the TMP-derived peptides might then be re-imported by TAP into the ER for MHC-I loading. This ERAD-driven antigen retrotranslocation might be facilitated by lipid bodies (LBs) \cite{Bougnères2009}, since LBs can serve as cytosolic sites for ubiquitination of ER-derived cargo (Fujimoto and Ohsaki, 2006). \geert{missen nog bijna alle citaties in dit stuk}

A second possibility is that TMPs are proteolytically processed by intramembrane proteases that cleave TMHs while they are still membrane embedded. Supporting this hypothesis is the well established notion that peptides generated by signal peptide peptidases (SPPs), an important class of intramembrane proteases that cleave TMH-like signal sequences, are presented on a specialized class of MHC-I called HLA-E (Oliveira, 2015). The loading of peptides generated by SPP onto MHC-I does not depend on the proteosome and TAP, possibly because the peptides are directly released into the lumen of the ER (Oliveira, 2015). However, this mechanism would not explain how multispanner polytopic membrane proteins can be processed for antigen presentation, because SPPs only cleave TMH-like signal sequences at the N-terminus of a protein. Nevertheless, the presentation of peptides with a high hydrophobicity index was shown to be independent on TAP as well (Lautscham, 2001), suggesting the TMH peptides might perhaps be released directly in the ER lumen by other intramembrane proteases. \geert{missen nog bijna alle citaties in dit stuk}

A third possibility is that peptide processing and MHC-loading occur in multivesicular bodies (MVBs) (Oliveira, 2015). TMPs can be routed from the plasma membrane and other organelles by vesicular trafficking to endosomes. Eventually, these TMPs can be sorted by the endosomal sorting complexes required for transport (ESCRT) pathway into luminal invaginations that pinch off from the limiting membrane and form intraluminal vesicles. This thus results in MVBs where the membrane proteins destined for degradation are located in intraluminal vesicles. Upon the fusion of MVBs with lysosomes, the entire intraluminal vesicles including the TMPs are degraded (Gruenberg,  2019). Via this mechanism, TMPs might well be processed for antigen presentation, particularly since the loading of MHC class II molecules is well understood to occur in MVBs (Kleijmeer 2001, Peters 1991, Zwart 2005). However, such processing of membrane proteins in MVBs for antigen presentation poses a problem, because complexes of HLA-DR with its antigen-loading chaperon HLA-DM were only observed on intraluminal vesicles, but not on the limiting membranes of MVBs (Zwart 2005), indicating that epitope loading of MHC-II also occurs at intraluminal vesicles. This observation hence raises the question how the intraluminal vesicles carrying the TMPs destined for antigen presentation can be selectively degraded, while the intraluminal vesicles carrying the MHC-II remain intact. A second problem is that phagosomes carrying internalized microbes lack intraluminal vesicles, and it is hence unclear how TMPs from these microbes would be routed to MVBs for MHC-II loading (Zwart et al 2005). \geert{check of correct zo, en weer referenties missen}.

Alternatively to the enzymatic degradation of lipids in MVBs by lipases (Sander 2016; Gilleron 2016), they might be oxidatively degraded by reactions with radical oxygen species (ROS) produced by the NADPH oxidase NOX2 (Dingjan 2016). This oxidation can result in a destabilization and disruption of membranes (Dingjan 2016) and might thereby lead to the extraction of TMPs. Due to the hydrophobic nature of TMHs, however, the extracted proteins will likely aggregate and it is unclear how these aggregates would be processed further for MHC loading. \geert{referenties missen}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Experimental evidence for MHC presentation of TMH-derived epitopes}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\geert{Deze gehele subsection kan wat mij betreft wel weg (of naar de 'minor discussion' in de supplement). Het is deels een herhaling van resultaten (1e en 3e paragraaf). DRiPs in paragraaf 2 kan echt niet alles verklaren, want MHC-II is opname van heel antigen. In parafraag 4, zal de methodological bias zal alleen maar resulteren in het onderschatten van presentatie van TMHs, en is dus ook niet echt relevant hier.}
% \paragraph{Over-presentation in vivo in MHC-I and MHC-II is robust
From reanalysis of peptide elution studies, we show that 1.4\% of epitopes that are naturally presented on MHC-I 
can be mapped to stem from a TMH. 
We also find that MHC-II presents epitopes derived from TMHs, 
as we find that 3.9-4.0\% of sequenced epitopes
can be uniquely traced onto a human reference proteome (also shown in table \ref{tab:elution}). \geert{zie eerdere vraag. Zijn deze percentages van totale peptides, of van humane peptides?}
